Data Migration under a Changing Schema in Ampersand

### 2023-11-19 2022Migration/articleMigrationFACS.tex
minor

   Before a migration starts we assume that the existing system is up and running and all of its invariants are satisfied.
   The ingress is directing traffic to the existing system.
   The data migration has three distinct steps:
\begin{description}
\item[Pre-deploy]
      The migration engineer deploys both the migration system and the desired system with their initial data,
      to allow both systems to satisfy their initial invariants before going live.
      The existing data is still in the existing system and the ingress still directs all traffic to the existing system.
      So, users notice no difference yet.
      The migration system starts copying data from the existing system.
\item[moment of transition (MoT)]
      After the migration system is done copying data, the migration engineer switches all traffic to the migration system
      and deploys the new functionality at the same time.
      This is the moment users will notice the difference.
      Even though the migration system may look like the desired system in the eyes of an average user,
      it relaxes the blocking invariants and keeps the system operational.
      Since the existing system receives no more traffic, its activity will come to a halt and its data will become static.
      The migration system stays active until all invariants of the desired system are satisfied and it can take over the work from the migration system.
\item[moment of completion (MoC)]
      Once the invariants of the desired system are satisfied, the migration engineer switches all traffic to the desired system.
      After that, she can safely remove both the migration system and the existing system, including their data.
\end{description}

@@

\section{Proof of Concept}
\label{sct:PoC}
   By way of proof of concept (PoC), we have built a migration system in Ampersand.
   To demonstrate it in the context of this paper, the existing system, $\pair{\dataset}{\schema}$, is rather trivial.
   It has no constraints and just one relation, $\declare{r}{A}{B}$.
   Its population is $A=\{a_1,a_2,a_3\}$, $B=\{b_1\}$, and $\pop{r}{\dataset}=\{\pair{a_1}{b_1}\}$.
   The schema of the migration system, $\schema_\migrsys$, follows from definition~\ref{eqn:schema migrsys}.
   Figure~\ref{fig:PoC} exhibits four successive screenshots,
   featuring $\overleftarrow{r}$ as {\small\verb#old_r#},
   $\overrightarrow{r}$ as {\small\verb#new_r#}.
\begin{figure}[bht]
   \begin{center}
      \makebox[\textwidth][c]{\includegraphics[scale=.07]{figures/screenshots.drawio.png}}%
   \end{center}
\caption{Four successive screenshots in the PoC}
\label{fig:PoC}
\end{figure}

>>>>

   Before a migration starts we assume that the existing system is up and running and all of its invariants are satisfied.
   The ingress is directing traffic to the existing system.
   The data migration has three distinct steps:
\begin{description}
\item[Pre-deploy]
      The migration engineer deploys both the migration system and the desired system with their initial data,
      to allow both systems to satisfy their initial invariants before going live.
      The existing data is still in the existing system and the ingress still directs all traffic to the existing system.
      So, users notice no difference yet.
      The migration system starts copying data from the existing system.
\item[moment of transition (MoT)]
      After the migration system is done copying data, the migration engineer switches all traffic to the migration system.
      This is the moment users will notice the difference because the traffic switch also deploys the functionality of the desired system.
      So, in the eyes of an average user, the migration system may look like the desired system.
      However, the migration system relaxes the blocking invariants of the desired system until users resolve their violations.
      Since the existing system receives no more traffic, its activity will come to a halt and its data will become static.
      The migration system stays active until all invariants of the desired system are satisfied and it can take over the work from the migration system.
\item[moment of completion (MoC)]
      Once the invariants of the desired system are satisfied, the migration engineer switches all traffic to the desired system.
      The blocking invariants of the desired system are now in effect, so users cannot violate them anymore.
      After this switch, the migration engineer can safely remove both the migration system and the existing system.
\end{description}

@@

\section{Proof of Concept}
\label{sct:PoC}
   By way of proof of concept (PoC), we have built a migration system in Ampersand.
   To demonstrate it in the context of this paper, the existing system, $\pair{\dataset}{\schema}$, is almost trivial.
   It has no constraints and just one relation, $\declare{r}{A}{B}$.
   Its population is $A=\{a_1,a_2,a_3\}$, $B=\{b_1\}$, and $\pop{r}{\dataset}=\{\pair{a_1}{b_1}\}$.
   The schema of the migration system, $\schema_\migrsys$, follows from definition~\ref{eqn:schema migrsys}.
   Figure~\ref{fig:PoC} exhibits four successive screenshots,
   featuring $\overleftarrow{r}$ as {\small\verb#old_r#},
   $\overrightarrow{r}$ as {\small\verb#new_r#}.
\begin{figure}[bht]
   \begin{center}
      \makebox[\textwidth][c]{\includegraphics[scale=.07]{figures/screenshots.drawio.png}}%
   \end{center}
\caption{Four successive screenshots in the PoC}
\label{fig:PoC}
\end{figure}
